Evaluation of probabilistic photometric redshift estimation approaches for \textsc{LSST}

### 2019-08-22 paper/tex/analysis.tex
address most remaining comments

\subsection{Metrics of an ensemble of \pzpdf s}
\label{sec:qualmet}
%(Rongpu Zhou, Jeffrey Newman, Alex Malz)
%(Alex Malz, Peter Freeman, Sam Schmidt)
%(John Soo, Rongpu Zhou)

@@

A quantile of a distribution is the value $q$ at which the CDF of the distribution is equal to $Q$; percentiles and quartiles are familiar examples of linearly spaced sets of 100 and 4 quantiles, respectively.
The quantile-quantile (QQ) plot serves as a graphical visualization for comparing two distributions, where the quantiles of one distribution are plotted against the quantiles of the other distribution, providing an easy way to qualitatively assess the consistency between an estimating distribution and a true distribution.
The closer the QQ plot is to diagonal, the closer the match between the distributions.
% In this paper, QQ plots are used for two purposes: (1) for comparing $N(z)$ from \pzpdf s (estimated using Eq.~\ref{eq:stacked}) with the true $N(z)$, and (2) for assessing the overall consistency of an ensemble of \pzpdf s with their true redshifts on a population level, where the distribution of the PIT values (see previous section) is compared to a uniform distribution between $0$ and $1$.
%\red{is this enough reason to include both?--SJS}
%\aim{Maybe we should explain why we expect the distribution of CDFs to be uniform.}\red{Note: this is just describing the PIT.  I'm going to swap the order of PIT to come before this, and then say we are evaluating the distribution of PIT values.  I will also add why we expect this to be uniform--SJS}

>>>>

\subsection{Metrics of \pzpdf ensembles}
\label{sec:qualmet}
%(Rongpu Zhou, Jeffrey Newman, Alex Malz)
%(Alex Malz, Peter Freeman, Sam Schmidt)
%(John Soo, Rongpu Zhou)

@@

A quantile of a distribution is the value $q$ at which the CDF of the distribution is equal to $Q$; percentiles and quartiles are familiar examples of linearly spaced sets of 100 and 4 quantiles, respectively.
The quantile-quantile (QQ) plot serves as a graphical visualization for comparing two distributions, where the quantiles of one distribution are plotted against the quantiles of the other distribution, providing an easy way to qualitatively assess the consistency between an estimated distribution and a true distribution.
The closer the QQ plot is to diagonal, the closer the match between the distributions.
% In this paper, QQ plots are used for two purposes: (1) for comparing $N(z)$ from \pzpdf s (estimated using Eq.~\ref{eq:stacked}) with the true $N(z)$, and (2) for assessing the overall consistency of an ensemble of \pzpdf s with their true redshifts on a population level, where the distribution of the PIT values (see previous section) is compared to a uniform distribution between $0$ and $1$.
%\red{is this enough reason to include both?--SJS}
%\aim{Maybe we should explain why we expect the distribution of CDFs to be uniform.}\red{Note: this is just describing the PIT.  I'm going to swap the order of PIT to come before this, and then say we are evaluating the distribution of PIT values.  I will also add why we expect this to be uniform--SJS}

### 2019-08-22 paper/tex/appendix.tex
address most remaining comments

Though alternatives exist \citep{Malz:chippr}, ``stacking'' according to
\begin{equation}
  \label{eq:stacked}
  \hat{N}^{H}(z) \equiv \frac{1}{N_{tot}}\ \sum_{i}^{N_{tot}}\ \hat{p}^{H}_{i}(z)
\end{equation}
is the most widely accepted method for obtaining $\hat{N}^{H}(z)$ as an estimator of the redshift distribution from \pzpdf s derived by a method $H$.
Though we do not endorse the use of the stacked estimator of the redshift distribution, we use it under the untested assumption that the response of our metrics of $\hat{N}^{H}(z)$ will be analogous to the same metrics applied to a principled estimator of the redshift distribution.

@@

\annz, \gpz, and \metaphor\ show signs of overtraining, estimating enhanced peaks and diminished troughs relative to the training set, an obstacle that may be overcome with adjustment of the implementation.

@@

Figure~\ref{fig:nz_stats} shows the quantitative Kolmogorov-Smirnoff (KS), Cramer-Von Mises (CvM), and Anderson Darling (AD) test statistics for each of the codes for the $\hat{N}(z)$ based measures.  The horizontal lines show the the result of a bootstrap resampling of the training set using 30,000 samples for \trainz.  As mentioned in Section~\ref{sec:pitqq} this represents a conservative idealized limit on expected performance for a modest-sized representative training set of galaxies.  The AD bootstrap statistic is elevated due to this statistic's sensitivity to distribution tails.
The stacked estimators of the redshift distribution for \cmnn\ and \trainz\ best estimate $\tilde{N}(z)$ under these metrics, and \bpz, \gpz, and \tpz\ are within a factor of two of the conservative limit for all statistics.  The only codes that do especially poorly are \eazy, \lephare, \metaphor, and \skynet.
It is unsurprising that \cmnn\ scores well, as with a near perfectly representative training set means that choosing neighbouring points in color/magnitude space should lead to excellent agreement in the final $\hat{N}(z)$ estimate.

>>>>

Though alternatives exist \citep{Malz:chippr}, ``stacking'' according to
\begin{equation}
  \label{eq:stacked}
  \hat{N}^{H}(z) \equiv \frac{1}{N_{tot}}\ \sum_{i}^{N_{tot}}\ \hat{p}^{H}_{i}(z)
\end{equation}
is the most widely accepted method for obtaining $\hat{N}^{H}(z)$ as an estimator of the redshift distribution from \pzpdf s derived by a method $H$.
Though the use of the stacked estimator of the redshift distribution is not formally correct, we use it under the untested assumption that the response of our metrics of $\hat{N}^{H}(z)$ will be analogous to the same metrics applied to a principled estimator of the redshift distribution.

@@

\annz, \gpz, and \metaphor\ show possible signs of overtraining: showing exaggerated peaks and troughs relative to the training set.  Further investigation on overtraining is needed, if present this is an obstacle that may be overcome with adjustment of the implementation.

@@

Figure~\ref{fig:nz_stats} shows the quantitative Kolmogorov-Smirnoff (KS), Cramer-Von Mises (CvM), and Anderson Darling (AD) test statistics for each of the codes for the $\hat{N}(z)$ based measures.  The horizontal lines show the the result of a bootstrap resampling of the training set using 30,000 samples for \trainz.  As mentioned in Section~\ref{sec:pitqq} this represents a conservative idealized limit on expected performance for a modest-sized representative training set of galaxies.  The AD bootstrap statistic is elevated due to this statistic's sensitivity to distribution tails.
The stacked estimators of the redshift distribution for \cmnn\ and \trainz\ best estimate $\tilde{N}(z)$ under these metrics, and \bpz, \gpz, and \tpz\ are within a factor of two of the conservative limit for all statistics.  The only codes that do especially poorly are \eazy, \lephare, \metaphor, and \skynet.
It is unsurprising that \cmnn\ scores well, as with a nearly complete and representative training set choosing neighbouring points in color/magnitude space to construct an estimator should lead to excellent agreement in the final $\hat{N}(z)$.

### 2019-08-22 paper/tex/data.tex
moving information from LePhare
address most remaining comments

Our mock catalogue is derived from the \textsc{Buzzard}-highres-v1.0  of De Rose et al., in prep, Wechsler et al., in prep) catalogue.
% UPDATE THIS CITATION NOW THAT THE PAPER IS OUT
\textsc{Buzzard} is built on a dark matter-only N-body simulation of $2048^{3}$ particles in a $400$ Mpc h$^{-1}$ box.
The lightcone was constructed from smoothing and interpolation between a set of time snapshots.
Dark matter halos were identified using the \texttt{Rockstar} software package \citep{Behroozi:13} and then populated with galaxies with a stellar mass and absolute $r$-band magnitude in the \sdss\ system determined using a sub-halo abundance matching model constrained to match both projected two-point galaxy clustering statistics and an observed conditional stellar mass function \citep{Reddick:13}.

To assign a spectrum to each galaxy, the Adding Density Dependent Spectral Energy Distributions (SEDs) procedure (\texttt{ADDSEDS}, deRose in prep.)\footnote{\url{https://github.com/vipasu/addseds}} was used.
\texttt{ADDSEDS} uses a sample of $\sim 5\times 10^{5}$ galaxies from the magnitude-limited \sdss\ Data Release 6 Value Added Galaxy Catalogue \citep{Blanton:05} to train an empirical relation between absolute $r$-band magnitude, local galaxy density, and SED.
Each \sdss\ spectrum is parameterized by five weights corresponding to a weighted sum of five basis SED components using the \texttt{k-correct \red{v4\_3?}} software package\footnote{\url{http://kcorrect.org}} \citep{Blanton:07}.

@@

Following the findings of \citet{Bernstein:10}, \citet{Masters:2017} that only $~10^{4}$ spectra are necessary to calibrate \pz s to Stage IV requirements, we aimed to set aside a randomly selected training set of $3-5\times 10^{4}$ galaxies, $\sim 10\%$ of the full sample.
After all cuts described above, we designated the \textit{DC1 training set} of $44\,404$ galaxies for which observed photometry, true SEDs, and true redshifts would be provided to all codes codes and the blinded \textit{DC1 test set} of $399\,356$ galaxies for which photometry alone would be provided to all codes and \pzpdf s would be requested.
%The resulting catalogues contain $111\,171$ training galaxies and $1\,000\,883$ test galaxies.

>>>>

Our mock catalogue is derived from the \textsc{Buzzard}-highres-v1.0 catalogue  \citep[][Wechsler et al., in prep]{DeRose:19}.
\textsc{Buzzard} is built on a dark matter-only N-body simulation of $2048^{3}$ particles in a $400$ Mpc h$^{-1}$ box.
The lightcone was constructed from smoothing and interpolation between a set of time snapshots.
Dark matter halos were identified using the \texttt{Rockstar} software package \citep{Behroozi:13} and then populated with galaxies with a stellar mass and absolute $r$-band magnitude in the \sdss\ system determined using a sub-halo abundance matching model constrained to match both projected two-point galaxy clustering statistics and an observed conditional stellar mass function \citep{Reddick:13}.

To assign a spectrum to each galaxy, the Adding Density Dependent Spectral Energy Distributions (SEDs) procedure (\texttt{ADDSEDS}, deRose in prep.)\footnote{\url{https://github.com/vipasu/addseds}} was used.
\texttt{ADDSEDS} uses a sample of $\sim 5\times 10^{5}$ galaxies from the magnitude-limited \sdss\ Data Release 6 Value Added Galaxy Catalogue \citep{Blanton:05} to train an empirical relation between absolute $r$-band magnitude, local galaxy density, and SED.
Each \sdss\ spectrum is parameterized by five weights corresponding to a weighted sum of five basis SED components using the \texttt{k-correct} software package\footnote{\url{http://kcorrect.org}} \citep{Blanton:07}.

@@

Following the findings of \citet{Bernstein:10}, \citet{Masters:2017} that only $~10^{4}$ spectra are necessary to calibrate \pz s to Stage IV requirements, we aimed to set aside a randomly selected training set of $3-5\times 10^{4}$ galaxies, $\sim 10\%$ of the full sample.
After all cuts described above, we designated the \textit{DC1 training set} of $44\,404$ galaxies for which observed photometry, true SEDs, and true redshifts would be provided to all codes codes and the blinded \textit{DC1 test set} of $399\,356$ galaxies for which photometry alone would be provided to all codes and \pzpdf s would be requested.
%The resulting catalogues contain $111\,171$ training galaxies and $1\,000\,883$ test galaxies.
The \lsst\ photometric filter transmission curves were also considered public information that could be used by any code.

### 2019-08-22 paper/tex/methods.tex
address most remaining comments

Here we summarize the twelve \pzpdf\ codes compared in this study, summarized in Table~\ref{tab:list_of_codes}, which include both established and emerging approaches in template fitting and machine learning.
Though not exhaustive, this sample represents codes for which there was sufficient expertise within the \lsstdesc\ Photometric Redshifts Working Group; the authors welcome interest from those outside \lsstdesc\ to have their codes assessed in future investigations that build upon this one.

@@

\subsection{Training-based Approaches}
\label{sec:trainingcodes}

@@

Here, we make two modifications to the implementation of \citet{Graham:17} to comply with the controlled experimental conditions.
First, we do not impose nondetections on galaxies fainter than the expected \lsst\ 10-year limiting magnitude or bright enough to saturate with \lsst's CCDs, instead using all of the photometry for the DC1 test and training sets.
Second, we apply the initial colour cut to the training set before calculating the Mahalanobis distance in order to accelerate processing and use a magnitude pseudo-prior as in \citet{Graham:17}, but for both we use cut-off values corresponding to the DC1 training set galaxies' colours and magnitudes.

@@

To smooth the long tail in the distribution of magnitude errors, we use the log of the magnitude errors, improving numerical stability and eliminating the need for constraints on the optimization process.
% We use principaphoto-\mathinhead{$z$}{z} interim posteriorsl component analysis (PCA) to decorrelate the data
% WHAT IS BEING DECORRELATED HERE?
Unobserved magnitudes $x_{\rm u} = \mu_{\rm u} + \Sigma_{\rm uo}\Sigma_{\rm oo}^{-1}(x_{\rm o} - \mu_{\rm o})$ were imputed from observed magnitudes $x_{\rm o}$ and the training set mean $\mu$ and covariance $\Sigma$ using a linear model.
This is the optimal expected value of the unobserved variables given the observed ones under the assumption that the distribution is jointly Gaussian; note that this reduces to a simple average if the covariates are independent with $\Sigma_{\rm uo} = 0$.
We reserved for validation 20\% of the training set and used the Variable Covariance option in \textsc{GPz} with 200 basis functions, neglecting to apply cost-sensitive learning options.

@@

In this work, we used a hierarchical KNN to replace nondetections with values based on their neighbors.
The usual cross-validation step was also omitted for this study.

@@

To run \tpz, we replaced nondetections with an approximation of the $1\sigma$ detection threshold based on the error forecast of the 10-year \lsst\ data, i.~e.~$dm = 2.5 \log (1 + N/S)$ where $dm \sim 0.7526$ mag for $N/S = 1$.
We calibrated \tpz\ with the Out-of-Bag cross-validation technique \citep{Breiman:84,Carrasco_Kind:13} to evaluate its predictive validity and determine the relative importance of the different input attributes.
We grew 100 trees to a minimum leaf size of 5 using the $ugri$ magnitudes, all $u-g, g-r, r-i, i-z, z-y$ colours, and the associated errors, as the $z$ and $y$ magnitudes did not show significant correlation with the redshift in our cross-validation.
We partitioned our redshift space into 200 bins and smoothed each individual PDF with a smoothing scale of twice the bin size.

>>>>

Here we summarize the twelve \pzpdf\ codes compared in this study, summarized in Table~\ref{tab:list_of_codes}, which include both established and emerging approaches in template fitting and machine learning.
Though not exhaustive, this sample represents codes for which there was sufficient expertise within the \lsstdesc\ Photometric Redshifts Working Group.  Some aspects of data treatment were left to the individual code runners, for example, whether/how to split the available data with known redshifts into separate training and validation sets. Another key difference is the treatment of non-detections in one or more bands: some codes choose to ignore a band, others replace the value with either an estimate for the detection limit, the mean of other values in the training set, or another default value. There are varying conventions among machine learning based codes for treatment of non-detections, and no one prescription dominates in the photo-z literature. The specific choices for each code affect the results, and contribute to the implicit prior influencing their output. However, we remind the reader that only 2.0 per cent of our sample has non-detections, almost exclusively in the u-band, and thus should not dominate the code performance differences.  The authors welcome interest from those outside \lsstdesc\ to have their codes assessed in future investigations that build upon this one.

@@

\subsection{Machine Learning-based Approaches}
\label{sec:trainingcodes}

@@

Here, we make two modifications to the implementation of \citet{Graham:17} to comply with the controlled experimental conditions.
First, we do not impose non-detections on galaxies fainter than the expected \lsst\ 10-year limiting magnitude or bright enough to saturate with \lsst's CCDs, instead using all of the photometry for the DC1 test and training sets.
Second, we apply the initial colour cut to the training set before calculating the Mahalanobis distance in order to accelerate processing and use a magnitude pseudo-prior as in \citet{Graham:17}, but for both we use cut-off values corresponding to the DC1 training set galaxies' colours and magnitudes.

@@

To smooth the long tail in the distribution of magnitude errors, we use the log of the magnitude errors, improving numerical stability and eliminating the need for constraints on the optimization process.
% We use principaphoto-\mathinhead{$z$}{z} interim posteriorsl component analysis (PCA) to decorrelate the data
% WHAT IS BEING DECORRELATED HERE?
Unobserved magnitudes $x_{\rm u} = \mu_{\rm u} + \Sigma_{\rm uo}\Sigma_{\rm oo}^{-1}(x_{\rm o} - \mu_{\rm o})$ were imputed from observed magnitudes $x_{\rm o}$ and the training set mean $\mu$ and covariance $\Sigma$ using a linear model.
This is the optimal expected value of the unobserved variables given the observed ones under the assumption that the distribution is jointly Gaussian; note that this reduces to a simple average if the covariates are independent with $\Sigma_{\rm uo} = 0$.
We reserved for validation 20\% of the training set and used the Variable Covariance option in \textsc{GPz} with 200 basis functions (see \citet{Almosallam:15b} for details), neglecting to apply cost-sensitive learning options.

@@

In this work, we used a hierarchical KNN to replace non-detections with values based on their neighbors.
The usual cross-validation of redshift estimates and PDFs was also omitted for this study.

@@

To run \tpz, we replaced non-detections with an approximation of the $1\sigma$ detection threshold based on the error forecast of the 10-year \lsst\ data, i.~e.~$dm = 2.5 \log (1 + N/S)$ where $dm \sim 0.7526$ magnitudes for $N/S = 1$.
We calibrated \tpz\ with the Out-of-Bag cross-validation technique \citep{Breiman:84,Carrasco_Kind:13} to evaluate its predictive validity and determine the relative importance of the different input attributes.
We grew 100 trees to a minimum leaf size of 5 using the $ugri$ magnitudes, all $u-g, g-r, r-i, i-z, z-y$ colours, and the associated errors, as the $z$ and $y$ magnitudes did not show significant correlation with the redshift in our cross-validation.
We partitioned our redshift space into 200 bins and smoothed each individual PDF with a smoothing scale of twice the bin size.
